A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-07 Paper/main.tex
update

>>>>

\textbf{Port Scanning attacks.}
According to~\cite{kanlayasiri2000rule}, Port scanning attack is a method for discovering exploitable communication channels that has been used for a long time. The key idea is to probe the network ports and then keep the information about them that are useful for an attack.
Hackers frequently employ a port scan approach to find gaps or weak spots in a network. Attackers can use a port scan attack to identify open ports and determine if they are accepting or rejecting data; additionally, it can show whether a company uses firewalls or other active security measures. 
Assuming that our server has multiple network inferfaces, thus multiple IP addresses; the victim machine can presume that port scans are malicious when the counter of reached ports overcomes a specific threshold in a short amount of time. The formulas are the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{NewPort}(\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S) = (i1=ip1)\land(i2=ip2)\land (port \notin S)\\
    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S,\underline{\text{int}}\ C) = (time-t<d)\\
    & \quad\quad\rightarrow((\mathtt{NewPort}(i1,i2,S)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i1,i2,S\cup\{port\},C)))  \\
    & \quad\quad\land(\neg \mathtt{NewPort}(i1,i2,S)\rightarrow\mathtt{Counter}(t,d,c,i1,i2,S,C))) \\
    & \underline{\text{mon}}\ \mathtt{PortScanSafe} =\ \mathtt{Always}(\mathtt{Counter}(time,100,1,ip1,ip2,\{port\},10)) 
\end{align*}
in this case we have two IPs; $ip1$ is the source IP, while $ip2$ is the destination one. The parameterized $\mathtt{Counter}$ rule states that the number of port scans between a source and destination IP  address pair $(i1,i2)$ recorded do not surpass a threshold $C$ within time $d$. The rule $\mathtt{NewPort}$ checks if the port number,
involved in any communication between the IP addresses $i1$ and $i2$, exists in the set $S$ of all port numbers (involved in all communications between $i1$ and $i2$) within the timeout period $d$.
